---
title: "class7"
author: "DongSu Kim"
date: "1/28/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Source
```{r}
source("http://tinyurl.com/rescale-R")
```


These are functons here
```{r}
rescale <- function(x, na.rm=TRUE, plot=FALSE, ...) {
  # Our rescale function from the end of lecture 9

  if(na.rm) {
    rng <-range(x, na.rm=TRUE)
  } else {
    rng <-range(x)
  }

  answer <- (x - rng[1]) / (rng[2] - rng[1])
  if(plot) { 
    plot(answer, ...) 
  }

  return(answer)
}
```
! = NOT

```{r}
rescale2 <- function(x, na.rm=TRUE, plot=FALSE, ...) {
 if( !is.numeric(x) ) {
 stop("Input x should be numeric", call.=FALSE)
 }
 rng <-range(x, na.rm=na.rm)

 answer <- (x - rng[1]) / (rng[2] - rng[1])
 if(plot) {
 plot(answer, ...)
 }
 return(answer)
}

```

new value
```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

is.na = helps found missing values -> pick out na 
```{r}
is.na(x)

typeof(is.na(x))
```

function that both there is no numeric value 
```{r}
is.na(x) & is.na(y)
```

so Function: 
```{r}
is_both_na <- function(x,y){
  is.na(x) & is.na(y)
}

is_both_na(x,y)
```

new problem when z -> function break

```{r}
z <- c(1,1)
is_both_na(x,z)
```

NEW FUNCTION CHANGE to stop the function
```{r}
is_both_na <- function(x,y){
  if(length(x) != length(z)){
    stop("your inputs aren't the same length")
  }
  is.na(x) & is.na(y)
}



```
Function : na.number = how many both true 
            na. which = where is the both true 
```{r}
rescale_new <- function(x, y) {
  ## Print some info on where NA's are as well as the number of them 
  if(length(x) != length(y)) {
    stop("Input x and y should be vectors of the same length", call.=FALSE)
  }
  
  na.in.both <- ( is.na(x) & is.na(y) )
  na.number <- sum(na.in.both)
  na.which <- which(na.in.both)
  
  message("Found ", na.number, " NA's at position(s):", 
          paste(na.which, collapse=", ") ) 
  
  return( list(number=na.number, which=na.which) )
}

rescale_new(x,y)

```

```{r}
x <- df1$IDs
y <- df2$IDs

intersect(x,y)

 ## what elements of x exists in y = TRUE 
x %in% y

  ## print out the subset of intersected as TRUE
x[x %in% y]

  ## how about y ?
y%in%x
y[y%in%x]

cbind(x[x%in%y],y[y%in%x])
```

Make a function!
```{r}
gene_intersect <- function(x, y) {
  cbind(x[x%in%y],y[y%in%x])
}


## modificaiton of the function with subset 
gene_intersect2 <- function(df1,df2) {
  cbind(df1[df1$IDs %in% df2$IDs,],
        df2[df2$IDs %in% df1$IDs,], "exp")
}




## modificaiton
colnames(df1) <- c("new_name", "exp")
colnames(df2) <- c("new_name", "exp")
colnames(df3) <- c("new-name", "exp")

gene_intersect3 <- function(df1,df2, gene.colname = "IDs") {
  cbind(df1[df1[,gene.colname] %in% df2[,gene.colname],],
        df2[df2[,gene.colname] %in% df1[,gene.colname],], "exp")
}

## modification SIMPLE AND CLEAR WAY!!!
gene_intersect4 <- function(df1,df2, gene.colname = "IDs") {
    df1.name <- df1[,gene.colname] #this here reurns a name
    df2.name <- df2[,gene.colname]
    
    df1.inds <- df1.name %in% df2.name
    df2.inds <- df2.name %in% df1.name
    
  cbind(df1[df1.inds,],
        df2[df2.inds,], "exp")
}

gene_intersect4(df1,df2,gene.colname = "new_name" )4
```

CRAN & BIOCONDUCTOR & ggplot2

##ggplot only takes dataframe file 

```{r}
library(ggplot2)
dim(iris) # a built-in data set
class(iris)
head(iris)
levels(iris$Species)
```
## Aestethic of ggpolot2 
how your data are represented viusally 
which data on the x, which data on the y

```{r}
myplot <- ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width)) #initialize a ggplot object 
class(myplot)
```

##Geometry
the genometic objects(pointsm lines, polygons) in the plot
shortcut function: geom point(), geom bar(), geom line()
```{r}
ggplot(data = iris, aes(x=Sepal.Length, y = Sepal.Width, color = Species))+ geom_point(aes(shape = Species), size=2) + geom_smooth(method = "lm")
```

```{r}
ggplot(iris, aes(Sepal.Length,fill=Species))+geom_bar(stat='count',binwidth = 0.3,alpha=0.7)
```


gg plot cheat sheet -> gives when to use the ggplot 