---
title: "Class: Structural Bioinformatics pt. 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB statistics

Here we inspect the types of structures in the main database for 3D biomlecular data - the PDB 

Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” >
“by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project
and
```{r}
#read in file
stats <- read.csv("Data Export Summary.csv", row.names = 1)
```

>Q1:Determine the percentage of structures solved by X-Ray and Electron Microscopy.

```{r}
#calculate percentage vlaues asked for 
ans <- stats$Total /sum(stats$Total) * 100
names(ans) <- rownames(stats)
round(ans, 2)

```


>Q2 Also can you determine what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers.

```{r}
round(sum(stats$Proteins) / sum(stats$Total) * 100,2)
```


>Q3: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

1289 of HIV-1 protease 


#Part 2 

Read a single PDB structue into R
```{r}
library(bio3d)

pdb <- read.pdb("1hsg")
pdb
```

```{r}
attributes(pdb)
```

```{r}
pdb$atom
```

##Select pretein only atoms
Here I will use the 'atom.select()' function to do this ... 
pdb <- READ.PDB("PBDID like '1HSG'")
Atom.select(pdb, other arguments ... )

```{r}
prt <- atom.select(pdb, "protein", value =TRUE)
lig <- atom.select(pdb,"ligand", value = TRUE)
```
And write out these new PDB onjects with the 'write.pdb()' function
```{r}
write.pdb(lig, file="1hsg_ligand.pdb")
write.pdb(prt, file="1hsg_protein.pdb")
```


