---
title: 'Class 06: R Functions'
author: "DongSu Kim"
date: "1/23/2020"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions 
## Level 2 heading
### Level 3 heading

Lets see more about **file import**(i.e. reading fi;es into R). The main read function in base R is `read.table()`

```{r}
t1 <- read.table("test1.txt")

```
We need to add arguments to get this file imported 
```{r}
t1 <- read.table("test1.txt", sep=",", header=TRUE)

```
Or I could just use `read.csv()` which ahs the arguments I want in this case!

```{r}
t1 <- read.csv("test1.txt")
```

```{r}
t2 <- read.table("test2.txt", header = TRUE, sep="$")
```
```{r}
t3 <- read.table("test3.txt", header = TRUE, sep ="")
```

#Back to functions

our first example function:

```{r}
add <- function(x, y=1) {
  #Sum the input x and y
  x + y
}

```

Let's use this function
```{r}
add(5)
add(7,3)
add(4,-2)
```
How does this work with vectors an input.
```{r}
add( c(1,2,4))
add( c(1,2,4),2)
add( c(1,2,4), c(4,2,1))
```

What is this `range()` function?
```{r}
x <- c(4,4,10,3,11)
max(x)
min(x)
range(x)
```
This is our second function:
```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}

rescale(x) 
rescale(1:10)

```
```{r}
# how can you fix this function?
rescale( c(1,2,NA,3,10) )
```
```{r}
x <- c(1,2,NA,3,10)
rng <- range(x)
rng
```
Find the problem and fix.
```{r}
x <- c(1,2,NA,3,10)
rng <- range(x, na.rm = TRUE)
rng
```
The final test for fixed one.
```{r}
rescale2 <- function(x) {
 rng <-range(x, na.rm = TRUE)
 (x - rng[1]) / (rng[2] - rng[1])
}

rescale2( c(1,2,NA,3,10))
```
Going too far!!
```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 print("is it me you are looking for?")
 if(plot) {
   print("Don't sing again please!")
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}
```
```{r}
rescale3(x, plot=FALSE)
```
```{r}
rescale3(x, plot=TRUE)
```

Hands on section 1B
```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b


plotb3(s1.b, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

```

>Q1. What type of onject is returned from `read.pdb()`

-> It is a list of 8 things and of class "pbd", "see"
```{r}
class(s1)

str(s1)
```

-> It is a list of amino acids 
```{r}
aa321(s1$seqres)
```


 >Q2 What does the trim.pdb() function do?
 
 ?trim.pdb
 
 >Q3. What input parameter would turn off the marginal black and grey rectangles in the
plots and what do they represent in this case?

sse=s1.chainA
 
 >Q4 what will be the better plot to compare different proteins?
 
```{r}
plotb3(s1.b, typ="l", ylab="Bfactor")
points(s2.b, typ="l", col="blue", lwd=2)
points(s3.b, typ="l", col="red", lwd=2)

```
>Q5  Which proteins are more similar to each other in their B-factor trends. How could
you quantify this? 

```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

 
>Q6 How would you generalize the original code above to work with any set of input
protein structures?

>Write your own function starting from the code above that analyzes protein drug interactions by reading in any protein PDB data and outputs a plot for the specified protein. (See class lecture 9 for further details).

```{r}

hw <- function(x,y,z) {
  
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b


plotb3(s1.b, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
}

hw(x,y,z)  
```

