---
title: "Class13: Structure based drug discovery"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download and clean HIV-Pr receptor structure 

Here we will download the 1HSG PDB structure from the online database,
remove the bound drug from Merk and write out a protein only and ligand only new set of PDB files. 

We will use the functions 'get.pdb()', 'read.pdb()','atom.select()' and finally 'write.pdb()'

```{r}
library(bio3d)

file.name <- get.pdb("1hsg")

pdb <- read.pdb(file.name)

pdb
```

Select Protein only
```{r}
prot <- atom.select(pdb,"protein",value=TRUE)
write.pdb(prot, file="1hsg_protein.pdb")
```

Select and write out ligand only file
```{r}
ligd <- atom.select(pdb,"ligand",value=TRUE)
write.pdb(ligd, file="1hsg_ligand.pdb")
```

ligd = MK1 files 
```{r}

ligd

```

pdbqt file will add the hydrognen atoms and the partial charge of each atoms in new column 

Read docking results

```{r}
res <-read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "result.pdb")
```

