---
title: "class10 02.06"
author: "DongSu Kim"
date: "2/6/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Analysis of Human Brest Cancer Cells

## 1. Exploratory data analysis

First we need to import (i.e. read) our input data. I downloaded the csv file from the class website: 

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```

There are some funkey things in this dataset that we will ignore for our analysis. This includes the first and second ID and Diagnosis columns and the funny last X column.

```{r}
# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[,3:32])
head(wisc.data)
```

> Q. How many patients do we have data for?

```{r}
nrow(wisc.data)
``` 

> Q. how many cancer and non-cancer?

```{r}
table(wisc.df$diagnosis)
```

> Q3. How many variables/features in the data are suffixed with _mean?

```{r}
grep("_mean", colnames(wisc.data), value =TRUE)
```
```{r}
length(grep("_mean", colnames(wisc.data)))
```



## 2. Principal Component Analysis

Before we do PCA we need to thnik or consider, wheter we should CALE our input. 

It is important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data include:

  - The input variables use different units of measurement.
  - The input variables have significantly different variances.
  
Check the mean and standard deviation of the features (i.e. columns) of the wisc.data to determine if the data should be scaled. Use the colMeans() and apply() functions like you’ve done before.

```{r}
# apply function => 1=row, 2=column mean= mean, sd= standard deviation
#round => round up to x number sig.fig

round(apply(wisc.data, 2, sd),2)
```


looks like we need to set scale=TRUE ! 

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp( wisc.data, scale=TRUE )
summary(wisc.pr)
#Culmulative Proportion is adding up of Proportion of Variance of PC1 and PC2 and so on... 
```
> Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

The 1st PC capture 44.27% of the original variance. Note that 72.6% in the first 3 PCs.  

>Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

>Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

Lets make some figures...
```{r}
biplot(wisc.pr)
```
This is a hot mess! We need to do our own PC1 vs PC2 plot and lets color by the diagnosis 

```{r}
attributes(wisc.pr)
```

```{r}
#m = red
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
abline(h=0, col="gray", lty=2)
abline(v=0, col="gray", lty=2)    
```

### Cluster in PC space 
##3. Hierarchical clustering
First lets see if we can cluster the original data 
Kmeans( x, centers = 2, nstart = 20)
HCLUST( dist(x))
 -> gives results of PCA to x 

```{r}
wisc.hc <- hclust(dist(wisc.data))
plot(wisc.hc)
```
This does not look good! Let's try and combine the result of PCA with clustering ... 

##5. Combining Methods 
Let’s see if PCA improves or degrades the performance of hierarchical clustering.

Using the minimum number of principal components required to describe at least 90% of the variability in the data, create a hierarchical clustering model with the linkage method="ward.D2". We use Ward’s criterion here because it is based on multidimensional variance like principal components analysis. Assign the results to wisc.pr.hclust.

```{r}
#input = wisc.pr$x[,1:3]
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:3]), method = "ward.D2")
plot(wisc.pr.hclust)
```

To get out clusters out of this tree we neeed to CUT it with the 'cuttree()' function. 

```{r}
grps3 <- cutree(wisc.pr.hclust, k=3)
table(grps3)
```
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps3)
```

We can use the 'table()' function to compare the #diagnosis vector with out cluster results vector. 

```{r}
table(grps3, wisc.df$diagnosis)
```

## 7. Prediction
We will use the 'predict()' function that will take our PCA model from before and new cancer cell data and project that data onto our PCA space.
```{r}
new <- read.csv("new_samples.csv")
new
```

Use the 'predict()' function with our previous PCA midel and new data 
```{r}
npc <- predict(wisc.pr, newdata=new)
npc
```

Now draw the PCA plot again and add our new data:
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
points(npc[,1], npc[,2], col="blue", pch=15
       , cex=2)
text(npc[,1], npc[,2], c(1,2), col="white")
```
> Q17. Which of these new patients should we prioritize for follow up based on your results? = patient 2 



